---
title: "IMED 7280 reshape demo"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reshape2)
```

This document pretty much has code only, so you can add your own notes to each section once you see what it does

NOTE: "gather" and "spread"  in tidyverse have similar BUT NOT IDENTICAL functions to melt and dcast - pick the one you want and play with it to get to know it better. 


```{r airquality}
summary(airquality)

```

Try the melt function

```{r}

airq.melt<- melt(airquality)
head(airq.melt)
summary(airq.melt)

```

Let's see what assigning ID variables does

```{r}
airq.melt2<- melt(airquality, id="Month")
head(airq.melt2)

```

What about multiple ID variables?

```{r}
airq.melt3<- melt(airquality, id=c("Month", "Day"))
head(airq.melt3)


```

dcast - the opposite of melt

```{r}

airq.cast<- dcast(airq.melt)

```

Need to add a formula LHS~RHS

```{r}
head(airq.melt)
airq.cast<- dcast(airq.melt, ~variable)
```

Why doesn't this work? hint: are there individual identifiers in this data??

```{r}
airq.melt4<- melt(cbind(id=rownames(airquality), airquality))
airq.cast<- dcast(airq.melt4, id~variable)

```

Homework: Try melting and recasting this in different forms. In what cases would you want the data in those forms??