---
title: "homework"
author: "Samantha Lee"
date: "23/03/2020"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Histogram

1.	Make a histogram of p values for CD4T cells
2.	Change default parameters to increase clarity
3.	Overlap histograms of p values for CD4T and Monocytes, and modify them to be clearer
4.	Repeat step 3 but use density plot instead, and modify as needed to make it clear
5.	ADVANCED: Make both an overlap of all the cell type density plots as well as a faceted version with one distribution per facet and decide which is best

Load data.  
```{r load_data}
load("~/Desktop/2019IMED7290-master/IMED7280_2019_datavis_example.rdata")
load("~/Desktop/2019IMED7290-master/IMED7280_2019_datavis_data.rdata")
```

Load libraries.  
```{r libraries}
library(tidyverse)
```

Make histogram of p-vals for CD4T cells
```{r pval_histo_cd4t}
#subset only cd4t data
ggplot(cord_adult_ct_sub %>% select(CD4T_P), aes(x=CD4T_P)) +
  geom_histogram(stat = "bin")
```

Make histogram of p-vals for CD4T cells for clear.
```{r pval_histo_cd4t_CLEAR}
#subset only cd4t data
#-log10 p vals to get better visualization
#change bin width to allow for refinement
#add in significance line
ggplot(cord_adult_ct_sub %>% select(CD4T_P), aes(x=-log10(CD4T_P))) +
  geom_histogram(stat = "bin", binwidth = 0.1) + 
  geom_vline(xintercept = -log10(0.05), linetype = "dashed") +
  theme_bw()
```

Overlap histograms of p values for CD4T and Monocytes, and modify them to be clearer
```{r overlap_histo}
#grab data
data <- cord_adult_ct_sub %>% select(CD4T_P,Mono_P)
data$cpg <- rownames(data)

#melt data to be in long format
data.melt <- reshape2::melt(data, id.vars="cpg", 
                            measure.vars=c("CD4T_P","Mono_P"),
                            variable.name="CellType",
                            value.name="Pval")

#now plot
ggplot(data.melt, aes(x=-log10(Pval), group=CellType, fill=CellType)) +
  geom_histogram(binwidth=0.1, alpha=0.5) +
  scale_fill_manual(values=c("turquoise","firebrick"),
                    labels=c("CD4T", "Monocytes")) +
  geom_vline(xintercept = -log10(0.05), linetype = "dashed") +
  theme_bw() +
  #i think the border is ugly so i removed it
  theme(panel.border = element_blank()) +
  theme(legend.position = c(0.8, 0.8), legend.background = element_blank())
```

Repeat step 3 but use density plot instead, and modify as needed to make it clear
```{r overlap_density}
#density plot
ggplot(data.melt, aes(x=-log10(Pval), group=CellType, fill=CellType)) +
  geom_density(alpha=0.5) +
  scale_fill_manual(values=c("turquoise","firebrick"),
                    labels=c("CD4T", "Monocytes")) +
  geom_vline(xintercept = -log10(0.05), linetype = "dashed") +
  theme_bw() +
  #i think the border is ugly so i removed it
  theme(panel.border = element_blank()) +
  theme(legend.position = c(0.8, 0.8), legend.background = element_blank())
```

ADVANCED: Make both an overlap of all the cell type density plots as well as a faceted version with one distribution per facet and decide which is best
```{r density_ALL}
#get just data for pvals
pvals <- cord_adult_ct_sub[,1:6]
#add it rownames for id
pvals$cpg <- rownames(pvals)
  
#change pvals to loooonnnggg format  
data.melt.all <- reshape2::melt(pvals, id.vars="cpg",
                            variable.name="CellType",
                            value.name="Pval")

ggplot(data.melt.all, aes(x=-log10(Pval), group=CellType, 
                          fill=CellType, col=CellType)) +
  geom_density(alpha=0.1) +
  scale_fill_brewer(palette = "Accent") +
  scale_color_brewer(palette = "Accent") +
  geom_vline(xintercept = -log10(0.05), linetype = "dashed") +
  theme_bw() +
  #i think the border is ugly so i removed it
  theme(panel.border = element_blank()) +
  theme(legend.position = c(0.8, 0.8), legend.background = element_blank())
```

```{r facet}
ggplot(data.melt.all, aes(x=-log10(Pval), group=CellType, 
                          fill=CellType, col=CellType)) +
  geom_density() +
  scale_fill_brewer(palette = "Accent") +
  scale_color_brewer(palette = "Accent") +
  geom_vline(xintercept = -log10(0.05), linetype = "dashed") +
  theme_bw() +
  #i think the border is ugly so i removed it
  theme(panel.border = element_blank()) +
  #theme(legend.position = c(0.8, 0.8), legend.background = element_blank()) +
  facet_wrap( ~ CellType, nrow=2) +
  theme(legend.position = "none")
```